---
title: "Homework - 2"
author: "Group 4"
date: "5/19/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
Perform principal component analysis on NHL.xlsx, which contains statistics of 30 teams in the National Hockey League. The description of the variables is provided in the ‘Description’ sheet of the file. Focus only on the variables 12 through 25, and create a new data frame.
• Input the new data frame to fa.parallel() function to determine the number of components to extract
• Input the new data frame to principal() function to extract the components. If raw data is input, the correlation matrix is automatically calculated by principal() function.
• Rotate the components
• Compute component scores
• Graph an orthogonal solution using factor.plot()
• Interpret the results

## First, import all the required libraries

``` {r message = FALSE}
library(dplyr)
library(readxl)
library(psych)
```


``` {r message = FALSE}
# Import the NHL excel file as a dataframe
NHL <- data.frame(read_xlsx("./data/NHL.xlsx", sheet = "Data"))

# Select the columns 13-26, the 1st column is the index column from excel
df <- NHL[, 13:26]

# Use Parallel Analysis Scree Plots to figure out the number of factors to extract
fa.parallel(df, fa = "pc", n.iter = 100, show.legend = FALSE)
```


From the above plot it will be appropriate to use 1 factor.

``` {r}
# Perform PCA with varimax orthogonal rotation
pc <- principal(df, nfactors = 2, rotate = "varimax", scores = TRUE)

pc
```

``` {r}

# Plot the components and analyze them
factor.plot(pc, choose = c(1), labels = colnames(df), title = "PCA Component 1")
factor.plot(pc, choose = c(2), labels = colnames(df), title = "PCA Component 2")

factor.plot(pc, labels = colnames(df))
```

Observations made from the plots - 
- Face and PPP loads only on Component-2. 
- gg, gag, five, PKP, sc1, tr1, lead1, lead2, wop and wosp load only on Component-1.
- shots and sag load on both Component-1 and Component-2.

---
