---
title: 'Homework #1'
author: "Group 4"
date: "5/10/2020"
output:
  pdf_document: default
  html_document: default
---

# First, let's import the required libraries
``` {r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(RColorBrewer)
```

## Problem 1

(a) Plot area vs.temp, area vs. month, area vs. DC, area vs. RH for January
through December combined in one graph. Hint: Place area on Y axis and use
2x2 matrix to place the plots adjacent to each other.

``` {r}
# Import the Dataframe
forestfires <- data.frame(read.csv("./data/forestfires.csv"), stringsAsFactors = FALSE)

# Convert the month column into factors and sort from Jan-Dec
forestfires$month <- factor(forestfires$month, 
                            levels = c("jan", "feb", "mar", 
                                       "apr", "may", "jun", 
                                       "jul", "aug", "sep", 
                                       "oct", "nov", "dec"))

# Create 4 scatter plots
p1 <- ggplot(forestfires, aes(temp, area)) + geom_point(color="#d63447") +
      ggtitle("Temp vs Area") +
      theme_classic()
p2 <- ggplot(forestfires, aes(month, area, color=month)) + geom_point() +
      scale_color_brewer(palette = "Set3") +
      theme_classic() +
      theme(legend.position = "none") +
      ggtitle("Month vs Area")
p3 <- ggplot(forestfires, aes(DC, area)) + geom_point(color="#d63447") +
      ggtitle("DC vs Area") +
      theme_classic()
p4 <- ggplot(forestfires, aes(RH, area)) + geom_point(color="#d63447") +
      ggtitle("RH vs Area") +
      theme_classic()
      
# Arrange plots P1-P4 into a 2x2 grid
fig <- ggarrange(p1, p2, p3, p4,
                 ncol=2, nrow=2)

plot(fig)

rm(list=c("p1", "p2", "p3", "p4", "fig"))
```

(b) Plot the histogram of wind speed (km/h).

``` {r}
# Create the Wind-Historgram
wind_hist <- ggplot(forestfires, aes(wind)) +
             geom_histogram(bins=10, fill="#08d9d6", color="black") +
             theme_classic() +
             ggtitle("Histogram of Wind Speed") +
             labs(x="Wind Speed (km/hour)")

# Plot 
plot(wind_hist)

rm(wind_hist)
```

(c) Compute the summery statistics (min, 1Q, mean, median, 3Q, max,) of part b.

``` {r}
quantiles <- quantile(forestfires$wind)

cat("Minimum Wind Speed is :", quantiles[[1]], "\n")
cat("1st Quantile of Wind Speed is :", quantiles[[2]], "\n")
cat("Mean Wind Speed is :", mean(forestfires$wind), "\n")
cat("Median Wind Speed is :", quantiles[[3]], "\n")
cat("3rd Quartile of Wind Speed is :", quantiles[[4]], "\n")
cat("Maximum Wind Speed is :", quantiles[[5]], "\n")

```

(d) Add a density line to the histogram in part b.

``` {r}
wind_hist_plus_density <- ggplot(forestfires, aes(x=wind, y=..density..)) +   
                          geom_histogram(bins=10, colour="black", fill="#08d9d6") + 
                          geom_density(aes(y=..density..), color="black") +
                          ylab("Density") + 
                          xlab("Wind Speed (km/hour)") + 
                          ggtitle("Histogram of Wind Speed") +
                          theme_bw()

plot(wind_hist_plus_density)

rm(wind_hist_plus_density)

```