---
title: "Homework-5"
author: "Group 4"
date: "6/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

The file ToyotaCorolla.xlsx contains the data on used cars (Toyota Corolla) on sale during late summer of 2004 in The Netherlands. It has 1436 records containing details on 38 attributes, including Price, Age, Kilometers, HP, and other specifications. The goal is to predict the price of a used Toyota Corolla based on its specifications.

Data Preprocessing: Create dummy variables for the categorical predictors (Fuel Type and Color). Split the data into training (50%), validation (30%), and test (20%) datasets.

``` {r}
library(dplyr)
library(readxl)
library(ggplot2)
library(caret)
library(tree)
library(rpart)
library(rpart.plot)
library(OneR)
library(reshape2)

ToyotaCorolla <- read_xlsx("ToyotaCorolla.xlsx", sheet = "data")

dummies <- dummyVars(~ Fuel_Type + Color, data = ToyotaCorolla, sep = ".")
dummies <-predict(dummies, ToyotaCorolla)

ToyotaCorolla <- cbind(select(ToyotaCorolla, -c("Fuel_Type", "Color")), 
                       as.data.frame(dummies))

set.seed(20)
split_sample <- sample(1:3, 
                       size = nrow(ToyotaCorolla), 
                       prob = c(0.50, 0.30, 0.20), 
                       replace = TRUE)

train_data <- ToyotaCorolla[split_sample == 1, ]
valididation_data <- ToyotaCorolla[split_sample == 2, ]
test_data <- ToyotaCorolla[split_sample == 3, ]

```

## (a) Run a regression tree (RT) with the output variable Price and input variables
Age_08_04, KM, Fuel_Type, HP, Automatic, Doors, Quarterly_Tax, Mfg_Guarantee,
Guarantee_Period, Airco, Automatic_Airco, CD Player, Powered_Windows, Sport_Model,
and Tow_Bar.

``` {r}
reg.tree <- rpart(Price ~ Age_08_04 + KM + Fuel_TypeCNG + Fuel_TypeDiesel + 
                     Fuel_TypePetrol + HP + Automatic + Doors + 
                     Quarterly_Tax + Mfr_Guarantee + Guarantee_Period + Airco +                               Automatic_airco + CD_Player + Powered_Windows + Sport_Model + 
                     Tow_Bar,
             data = train_data,
             method = "anova")

rpart.plot(reg.tree)

plotcp(reg.tree)

print(reg.tree$variable.importance)
```
(i) Which appear to be the three or four most important car specifications for
predicting the car’s price?

- The most important car specifications for predicting the car's price are Age_08_04, KM and Automatic_airco.

(ii) Compare the prediction errors of the training, validation, and test sets by
examining their RMS error and by plotting the three boxplots. What is happening
with the training set predictions? How does the predictive performance of the test
set compare to the other two? Why does this occur?

``` {r message = FALSE, warning = FALSE}
train_preds <- predict(reg.tree, train_data)
validation_preds <- predict(reg.tree, valididation_data)
test_preds <- predict(reg.tree, test_data)

train_error <- RMSE(train_preds, train_data$Price)
validation_error <- RMSE(validation_preds, valididation_data$Price)
test_error <- RMSE(test_preds, test_data$Price)

cat("Train Data RMSE", train_error, "\n")
cat("Validation Data RMSE", validation_error, "\n")
cat("Test Data RMSE", test_error, "\n")


df <- melt(as.data.frame(cbind(train_preds, validation_preds, test_preds)))
ggplot(data = df, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  ggtitle("Price Predictions Box Plot") +
  labs(x = "", y = "Price")
```
We can see from the RMSE error that the model performs the best on the train_data but that is to be expected. 
It performs slightly worse on the validation data. It performs the worst on the test data.

We can see that the training set predictions are similar to the actual Price values of the train data with a mean of approx $10000. There are 2 samples which are outliers in all three of the sets. 


(iv) If we used the full tree instead of the best pruned tree to score the validation set, how would this affect the predictive performance for the validation set? (Hint: Does the full tree use the validation data?)

``` {r}
full.tree <- rpart(Price ~ Age_08_04 + KM + Fuel_TypeCNG + Fuel_TypeDiesel + 
                     Fuel_TypePetrol + HP + Automatic + Doors + 
                     Quarterly_Tax + Mfr_Guarantee + Guarantee_Period + Airco +                               Automatic_airco + CD_Player + Powered_Windows + Sport_Model + 
                     Tow_Bar,
             data = train_data,
             method = "anova",
             control = list(cp = 0))

rpart.plot(full.tree)

plotcp(full.tree)
```

``` {r}
train_preds <- predict(full.tree, train_data)
validation_preds <- predict(full.tree, valididation_data)
test_preds <- predict(full.tree, test_data)

train_error <- RMSE(train_preds, train_data$Price)
validation_error <- RMSE(validation_preds, valididation_data$Price)
test_error <- RMSE(test_preds, test_data$Price)

cat("Train Data RMSE", train_error, "\n")
cat("Validation Data RMSE", validation_error, "\n")
cat("Test Data RMSE", test_error, "\n")
```

Using the full tree, we significantly reduce the RMSE error on both the validation and the test sets. Although the full tree doesn't use the validation data, it has more decision boundaries which are correctly separating the validation data hence reducing its error.

(b) Let us see the effect of turning the price variable into a categorical variable. First, create a new variable that categorizes price into 20 bins of equal counts. Now
repartition the data keeping Binned Price instead of Price. Run a classification tree
(CT) with the same set of input variables as in the RT, and with Binned Price as the
output variable. 


``` {r}
ToyotaCorolla$Price <- cut(ToyotaCorolla$Price, breaks = seq(4300, 32500, by = 1410))

set.seed(20)
split_sample <- sample(1:3, 
                       size = nrow(ToyotaCorolla), 
                       prob = c(0.50, 0.30, 0.20), 
                       replace = TRUE)

train_data <- ToyotaCorolla[split_sample == 1, ]
valididation_data <- ToyotaCorolla[split_sample == 2, ]
test_data <- ToyotaCorolla[split_sample == 3, ]
```

``` {r}
class.tree <- rpart(Price ~ Age_08_04 + KM + Fuel_TypeCNG + Fuel_TypeDiesel + 
                     Fuel_TypePetrol + HP + Automatic + Doors + 
                     Quarterly_Tax + Mfr_Guarantee + Guarantee_Period + Airco +                               Automatic_airco + CD_Player + Powered_Windows + Sport_Model + 
                     Tow_Bar,
             data = train_data)

rpart.plot(class.tree)

plotcp(class.tree)

print(class.tree$variable.importance)

```

(i) Compare the tree generated by the CT with the one generated by the RT. Are they
different? (Look at structure, the top predictors, size of tree, etc.) Why?

- The top predictors of both the classification and regression trees remain more or less the same except of Quarterly_Tax which is only important in the regression tree.

- The classification tree is both structurally more complex as well bigger than the regression tree. This difference is created because in a regression tree, the response value is the mean of all the other values in the particular region however in the classification tree, the response is the majority class of the region.

- The difference in shape is so apparent because decision trees are sensitive to change in training data.

(ii) Predict the price, using the RT and the CT, of a used Toyota Corolla with the
specifications listed in Table below.

``` {r}
df <- data.frame(Age_08_04 = 77,
                 KM = 117000,
                 Fuel_TypePetrol = 1,
                 Fuel_TypeDiesel = 0,
                 Fuel_TypeCNG = 0,
                 HP = 110,
                 Automatic = 0,
                 Doors = 5,
                 Quarterly_Tax = 100,
                 Mfr_Guarantee = 0,
                 Guarantee_Period = 3,
                 Airco = 1,
                 Automatic_airco = 0,
                 CD_Player = 0,
                 Powered_Windows = 0,
                 Sport_Model = 0,
                 Tow_Bar = 1)


df.pred.reg <- predict(reg.tree, df)
df.pred.class <- as.data.frame(predict(class.tree, df))
df.pred.class <- attributes(which.max(df.pred.class))$names

cat("Prediction of Regression Tree:", df.pred.reg, "\n")
cat("Prediction of Classification Tree:", df.pred.class, "\n")
```

(iii) Compare the predictions in terms of the predictors that were used, the magnitude
of the difference between the two predictions, and the advantages and
disadvantages of the two methods.

- The Regression Tree predicts the new data Price to be `$`7906.141. While the classification tree predicts that the price will be in the category/bin/range `$`7120-`$`8530.

- The advantage of using regression trees is that we can obtain an exact numerical prediction of our data.

- However classification trees can prove to be more useful in predicting numerical data in situations where we are more concerned with the range estimation of the data rather than it's average value.

---



